---
title: "Getting Started with shinyTAG"
author:
  - name: John Abbey
    affiliation: University of Zurich
date: "`r Sys.Date()`"
abstract: >
  shinyTAG provides an interactive Shiny web application for visualising
  MALDI Mass Spectrometry Imaging (MALDI-MSI) data processed by the gutenTAG
  package. This vignette describes how to launch the application, load data,
  and navigate the three main panels: the ion image viewer, the interactive
  spectrum panel, and the quality control dashboard.
output:
  BiocStyle::html_document:
    toc: true
    toc_float: true
vignette: >
  %\VignetteIndexEntry{Getting Started with shinyTAG}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, comment = "#>")
```

# Installation

Install shinyTAG from GitHub during development, or from Bioconductor once
available:

```{r install, eval=FALSE}
# Development version
devtools::install_github("JohnAbbey98/shinyTAG")
```

# Quick Start

Launch the application with a single call:

```{r launch, eval=FALSE}
library(shinyTAG)
runShinyTAG()
```

This opens the shinyTAG interface in your default browser.

# User Interface Overview

The interface consists of three main areas:

## Data Import Sidebar

Located on the left, the sidebar accepts two inputs:

- **Load .imzML file** — the raw or pre-processed MALDI-MSI dataset.
- **Load panel .csv** — the target mass panel, as loaded by
  `gutenTAG::readPanel()`. This CSV must contain at minimum `Name` and
  `FeatureMass` columns.

After selecting both files, click **Process** to run the full gutenTAG
pipeline: `readMSIData` → `preProcess` → `peakDetection` →
`generateMetapeaks` → `assignMetapeaks`. A status message confirms when
processing is complete.

> **Note on large files:** Processing files upwards of 10 GB may take several
> minutes. The status message will update once the pipeline finishes.

## Ion Image Panel

The top card displays a single-channel spatial ion image rendered by
`gutenTAG::imageChannel()`. Use the **Colour palette** dropdown to switch
between viridis colour scales (viridis, magma, plasma, inferno, cividis).

The displayed channel updates automatically when you click on a metapeak in
the spectrum panel below.

## Spectrum Panel

The interactive spectrum shows the peak count histogram, smoothed counts,
metapeak boundaries, panel feature masses, and seed m/z values — equivalent
to `gutenTAG::plotMetapeaks(interactive = TRUE)`.

**Click anywhere within a metapeak boundary** (red shaded region) to select
that channel. The ion image above will update immediately.

## Quality Control Dashboard

The QC card contains six tabs, each showing a static plot from gutenTAG:

| Tab | Function |
|-----|----------|
| Intensity Distribution | `plotIntensityDistribution()` |
| Mean-Variance | `plotMeanVariance()` |
| Geary's C | `plotGearysC()` |
| SNR | `plotSNR()` |
| QC Overview | `plotQCOverview()` |
| TIC Spatial | `plotTICSpatial()` |

Tabs are rendered on demand; only the active tab's plot is computed.

# Example Data

gutenTAG ships with a small example dataset that can be used to test the
application:

```{r example-paths, eval=FALSE}
imzml_path <- system.file("extdata/Example_data.imzML", package = "gutenTAG")
panel_path  <- system.file("extdata/ref_list.csv",       package = "gutenTAG")
```

These paths can be entered directly into the file upload inputs (or used
programmatically in tests).

# Session Information

```{r session-info}
sessionInfo()
```
